Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications

### 2024-05-21 index.Rmd
adding html run code

---
title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
author:
  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
  - César Cárdenas^[Instituto Antártico Chileno, Chile]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
always_allow_html: yes
output:
  pdf_document
---

```{=tex}
\fontsize{12}{16}
\selectfont{}
```
\newpage

```{=latex}
\setcounter{tocdepth}{3}
\tableofcontents
```

\newpage

@@

# ABSTRACT

The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity. The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill. By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics. This nuanced approach allows for a more refined understanding of different zones shape krill abundance, fishing mortality and recruitment patterns across the Antarctic Peninsula. Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

*Keywords: Krill populations, dynamic population, stock assessment, SS3, spatial implicit, management, CCAMLR.*

# INTRODUCTION

**(add refs)**

Antarctica, with its pristine and unique ecosystems, hosts one of the most crucial marine resources, the Antarctic krill (Euphausia superba). These small crustaceans play a pivotal role in the Southern Ocean's food web, serving as a primary food source for a wide range of marine species, including whales, seals, and seabirds. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.

In this context, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) plays a central role in the conservation and sustainable management of marine resources in the Southern Ocean. As part of its mandate, CCAMLR regularly conducts stock assessments to evaluate the status of key species, including Antarctic krill, and inform decision-making processes regarding fisheries management and conservation measures.

Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill ecology, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.

In response to this need, we propose a comprehensive stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach. Unlike previous assessments, our model will integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics. Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions, sea ice dynamics, and predator foraging behavior. By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies.

Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics. Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds. Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem. By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.

The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill. By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model. This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.

In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula. By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR. Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.

# METHODOLOGY

## Study Area in the Antarctic Peninsula

The study area for the assessment of krill populations in the Antarctic Peninsula is divided into strata to account for spatial heterogeneity. The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.

The strata in the Antarctic Peninsula may include different regions such as coastal areas, offshore waters, and areas influenced by major oceanographic features like currents and frontal systems. Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability.

![Study Area 48.1 and strata](Figs/481.png)

## Data Sources and Recolection

Stock Synthesis relies on multiple sources of data to inform the population dynamics model. These data sources may include:

1. **Fisheries Data:** Information on krill catches and fishing effort from commercial fisheries operating in the Antarctic Peninsula region.

2. **Population Monitoring Data:** Data collected from scientific research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.

3. **Environmental Data:** Environmental data such as sea surface temperature, chlorophyll concentration, and sea ice extent are also incorporated into the assessment. These data help to understand the influence of environmental factors on krill populations and their distribution.

4. **Predator information:** MAPPPD. Predator “fleet” to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.

In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment. These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.

## Conceptual Model

## Phases in model construction

The process of modeling and statistical analysis of a database can be structured according to the following guidelines:

- Contextualization of the problem. Definition of objectives and variables.
- Experiment design and data collection.
- Recording and preliminary processing of available information.
- Graphical inspection and trend identification.
- Consideration of distributional and relational hypotheses. Proposal of modeling.
- Model adjustment. Comparison and selection of the best model.
- Diagnosis and validation of the adjusted model.
- Assessment of the predictive capacity of the model and prediction.
- Interpretation and conclusions.

## Population Dynamics Model

@@

A typical state-space model for krill population dynamics can be represented as:

\[
N_t = N_{t-1} \cdot e^{(r - M)} + R
\]

Where:
- \(N_t\) is the abundance of krill at time \(t\).
- \(N_{t-1}\) is the abundance of krill at the previous time step.
- \(r\) is the intrinsic growth rate of the population.
- \(M\) is the mortality rate.
- \(R\) is the recruitment of new individuals into the population.

This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.

@@

The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21). SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*. SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data. The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].

@@

@@

@@

```{r eval=TRUE}
SSplotRunstest(base.model01, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

@@

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
```
```{r echo=FALSE}
ggarrange(retro01,
          retro1,
          retro2,
          retro3,
          retro4,
          retro5,
          retro6,
          retro7, ncol = 4, nrow = 4)
```

@@

\newpage
# DISCUSION

\newpage
# CONCLUSION

>>>>

---
title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---

@@

# OVERVIEW

@@

@@

The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21). SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*. SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data. The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* and *ss3diags* library [@Taylor2019; @Winker2023].

@@

## Run Models

@@

Read outputs

@@

```{r eval=FALSE}
SSplotRunstest(base.model01, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

@@

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

```{r echo=FALSE}
ggarrange(retro01,
          retro1,
          retro2,
          retro3,
          retro4,
          retro5,
          retro6,
          retro7, ncol = 4, nrow = 4)
```

@@
